A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-01-24 paper.tex
Cleaned up and commented bounded algorithm
Work on theory sections, copied over Leonid's proof, needs editing still

\documentclass{llncs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{tikz}
\usetikzlibrary{trees}

\algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
\algnewcommand{\EndIIf}{}

@@

\title{Untitled CAV Paper}

\author{Alexander Legg\inst{1} 
    \and Leonid Ryzhyk\inst{2}
    \and Nina Narodytska\inst{2}}

\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
    Communications and the Digital Economy and the Australian Research Council through the ICT
    Centre of Excellence program.} and UNSW \\
    \email{alexander.legg@nicta.com.au}
    \and Samsung Research}

@@

    Reactive synthesis techniques based on image computations have been shown
    to work well in many cases but suffer from state explosion in others.
    A different approach applies SAT solvers in a counterexample guided framework but
    is limited to bounded synthesis.  We present an extension of this technique
    to unbounded synthesis by employing interpolation.  Experimental results
    show this to be a promising alternative that solves some previously
    intractable instances.

@@

\begin{algorithm}
    \caption{Tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}
        \Function{treeFormula}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{ $\lnot \Call{E}{x^{gt}}$ }
        \Else
        \State \Return{ $\lnot \Call{E}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
        \EndIf
        \EndFunction
    \end{algorithmic}

    \begin{algorithmic}
        \Function{$\overline{treeFormula}$}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{\Call{E}{$x^{gt}$}}
        \Else
        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n})$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

The bounded synthesis algorithm described in \cite{narodytska2014} begins with
an empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing
on behalf of the environment because it chooses the first move in each step.
The empty game is passed to the SAT solver, which searches for a candidate
environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
then it is checked for a spoiling strategy by solving for the controller.  If
no spoiling strategy exists, that means our candidate is a winning strategy and
the algorithm terminates.  Otherwise we find a counterexample (Figure
\ref{fig:cex}), which is used to refine the empty game tree to include the
first move from the controller's spoiling strategy (Figure \ref{fig:refined}.
The algorithm continues by finding a new candidate for the environment (Figure
\ref{fig:candidate2}), and a new counterexample to refine the game abstraction
once again (Figure \ref{fig:refined2}).

@@

\begin{algorithm}
    \begin{algorithmic}
        \Function{solveAbstract}{$\langle s, r \rangle , gt$}
        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, gt$}
        \IIf{$r = n - 1$} \Return $cand$ \EndIIf 
        \State $gt' \gets gt$
        \Loop
            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf
            \State $cex \gets $ \Call{verify}{$\langle s, r \rangle, gt, cand$}
            \IIf{$cex = NULL$} \Return $cand$ \EndIIf
            \State $gt' \gets gt' \cup cex$
%%%            \State $gt' \gets \Call{refine}{gt', cex}$
            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, gt'$}
        \EndLoop
        \EndFunction
    \end{algorithmic}

%%%    \begin{algorithmic}
%%%        \Function{refine}{$gt, \langle leaf, spoiling \rangle$}
%%%            \State \Return $gt \cup \{leaf, spoiling\}$
%%%        \EndFunction
%%%    \end{algorithmic}

    \begin{algorithmic}
        \Function{findCandidate}{$\langle s, r \rangle, T$}
            \State $\hat{T} \gets \Call{gtExtend}{T}$
            \State $fml \gets $ \Call{treeFormula}{$\hat{T}$}
            \State $sol \gets $ \Call{SAT}{$s(X_T) \land fml$}
        \EndFunction
    \end{algorithmic}

    \begin{algorithmic}
        \Function{verify}{$\langle r, s\rangle, gt, cand$}
            \For{$l \in leaves(gt)$}
                \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle r, s\rangle, gt, l$}
                \State $spoiling \gets $ \Call{solveAbstract}{$\langle r', s' \rangle, \emptyset$}
            \EndFor
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

@@

\begin{algorithm}
    \caption{Amended tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}
        \Function{treeFormula}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}}$ }
        \Else
        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
        \EndIf
        \EndFunction
    \end{algorithmic}

    \begin{algorithmic}
        \Function{$\overline{treeFormula}$}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{\Call{E}{$x^{gt}$}}
        \Else
        \State \Return{ $\Call{$L^e[rank(gt)]$}{x^{gt}} \land$ \\
            $(\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n}))$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

This is the foundation for a recursive algorithm that consumes an entire tree
by removing a single step on each iteration. All learned states from which the
controller must lose are recorded in a set $L^c$.  This algorithm can also be
performed on environment-losing trees with the caveat that any state learnt at
a node of rank $n$ is only known to lose at ranks less than or equal to $n$.

@@

\subsection{Algorithm}

>>>>

\documentclass{llncs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{fixltx2e}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{tikz}
\usetikzlibrary{trees}

\algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
\algnewcommand{\EndIIf}{}
\MakeRobust{\Call}
\newcommand{\True}{\texttt{true}}
\newcommand{\False}{\texttt{false}}
\newcommand{\II}{\mathcal{I}}
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}

@@

\title{A SAT-Based Counterexample Guided Method for Unbounded Synthesis}

\author{Alexander Legg\inst{1} 
    \and Nina Narodytska\inst{2}
    \and Leonid Ryzhyk\inst{2}}

\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
    Communications and the Digital Economy and the Australian Research Council through the ICT
    Centre of Excellence program.} and UNSW \\
    \email{alexander.legg@nicta.com.au}
    \and Samsung Research America}

@@

    Reactive synthesis techniques based on constructing the winning region of
    the system have been shown to work well in many cases but suffer from state
    explosion in others.  A different approach, proposed recently, applies SAT
    solvers in a counterexample guided framework to solve the synthesis
    problem.  However, this method is limited to synthesising systems that
    execute for a bounded number of steps and is incomplete for synthesis with
    unbounded safety and reachability objectives.  We present an extension of
    this technique to unbounded synthesis.  Our method applies Craig
    interpolation to abstract game trees produced by counterexample-guided
    search in order to construct a monotonic sequence of may-losing regions.
    Experimental results based on SYNTCOMP 2015 competition benchmarks show
    this to be a promising alternative that solves some previously intractable
    instances.

@@

\begin{algorithm}
    \caption{Tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{ $\lnot \Call{E}{x^{gt}}$ }
        \Else
        \State \Return{ $\lnot \Call{E}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
        \EndIf
        \EndFunction
        \algstore{tf1}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{tf1}
        \Function{\textoverline{treeFormula}}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{\Call{E}{$x^{gt}$}}
        \Else
        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

This technique was proposed by Narodytska et al \cite{narodytska2014}. In
Section \ref{sect:unbounded} we propose an extension to synthesis of unbounded
games.

The bounded synthesis algorithm begins with an empty game as seen in Figure
\ref{fig:emptytree}.  Initially we are playing on behalf of the environment
because it chooses the first move in each step.  The empty game is passed to
the SAT solver, which searches for a candidate environment strategy.  If a
candidate is found (Figure \ref{fig:candidate}) then it is checked for a
spoiling strategy by solving for the controller.  If no spoiling strategy
exists, that means our candidate is a winning strategy and the algorithm
terminates.  Otherwise we find a counterexample (Figure \ref{fig:cex}), which
is used to refine the empty game tree to include the first move from the
controller's spoiling strategy (Figure \ref{fig:refined}.  The algorithm
continues by finding a new candidate for the environment (Figure
\ref{fig:candidate2}), and a new counterexample to refine the game abstraction
once again (Figure \ref{fig:refined2}). The full algorithm is listed in
Algorithm \ref{alg:bounded}.

@@

\begin{algorithm}
    \begin{algorithmic}[1]
        \Function{solveAbstract}{$\langle s, r \rangle , T$}
        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, T$} \Comment{Look for a candidate}
        \IIf{$r = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
        \State $T' \gets T$
        \Loop
            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, r \rangle, T, cand$} \Comment{Verify candidate}
            \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, T'$} \Comment{Solve refined AGT}
        \EndLoop
        \EndFunction
        \algstore{b1}
    \end{algorithmic}

    \begin{algorithmic}
        \algrestore{b1}
        \Function{findCandidate}{$\langle s, r \rangle, T$}
            \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
            \IIf{$sol = \texttt{unsat}$} 
                \Return $\emptyset$ \Comment{No candidate exists}
            \EndIIf
            \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
        \EndFunction
        \algstore{b2}
    \end{algorithmic}

    \begin{algorithmic}
        \algrestore{b2}
        \Function{verify}{$\langle s, r\rangle, gt, cand$}
            \For{$l \in leaves(gt)$}
            \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle s, r\rangle, gt, l$} \Comment{Get rank and state at leaf}
                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle r', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
                \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
            \EndFor
            \State \Return $\langle \True, \emptyset, \emptyset \rangle$
        \EndFunction
    \end{algorithmic}

    \caption{Bounded Synthesis}
    \label{alg:bounded}
\end{algorithm}

@@

\label{sect:unbounded}

@@

\begin{algorithm}
    \begin{algorithmic}[1]
        \Require $\sigma(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
        \Require \emph{Must-invariant} holds
        \Ensure \emph{Must-invariant} holds
        \Ensure $\sigma(X_T) \land B^M \equiv \bot$
        \Function{learn}{$\sigma, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with max rank
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
            \State $B^M \gets B^M \lor \mathcal{I}$
            \State \Call{learn}{$\sigma, T_1$}
        \EndFunction
        \algstore{learn}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{learn}
        \Require $\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
        \Require \emph{May-invariant} holds
        \Ensure \emph{May-invariant} holds
        \Ensure $\sigma(X_T) \land B^m[$\Call{rank}{$T$}$] \equiv \bot$
        \Function{\textoverline{learn}}{$\sigma, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with max rank
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
            \For{$i = 1$ to \Call{rank}{$n$}}
                \State $B^m[i] \gets B^m[i] \land \lnot \mathcal{I}$
            \EndFor
            \State \Call{\textoverline{learn}}{$\sigma, T_1$}
        \EndFunction
    \end{algorithmic}
    \caption{Learning algorithms}
    \label{alg:learn}
\end{algorithm}

\begin{algorithm}
    \caption{Amended tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{ $\lnot \Call{$B^M$}{x^{gt}}$ }
        \Else
        \State \Return{ $\lnot \Call{$B^M$}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
        \EndIf
        \EndFunction
        \algstore{tf2}
    \end{algorithmic}

    \begin{algorithmic}
        \algrestore{tf2}
        \Function{\textoverline{treeFormula}}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{\Call{E}{$x^{gt}$}}
        \Else
        \State \Return{ $\Call{$B^m[rank(gt)]$}{x^{gt}} \land$ \\
            $(\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n}))$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

This is the foundation for a recursive algorithm that consumes an entire tree
by removing a single step on each iteration (Algorithm \ref{alg:learn}). All
learned states from which the controller \emph{must} lose are recorded in a set
of \emph{bad} states $B^M$.  This algorithm can also be performed on
environment-losing trees with the caveat that any state learnt at a node of
rank $n$ is only known to lose at ranks less than or equal to $n$. We negate
the interpolants we learn this way and record them in a mapping of ranks to
sets of states $B^m$ which \emph{may} be bad states for the controller.

\begin{algorithm}
    \begin{algorithmic}[1]
        \Function{solveUnbounded}{$T$}
            \State $B^M \gets E$
            \State $B^m[0] \gets E$
            \For{$k = 1 \dots$}
                \IIf{\Call{SAT}{$I \land B^M$}}
                    \Return \texttt{unrealisable} \Comment{Losing in the initial state}
                \EndIIf
                \IIf{$\exists i \  B^m[i] \equiv B^m[i+1]$}
                    \Return \texttt{realisable} \Comment{Reached fixed point}
                \EndIIf
                \State $B^m[k] \gets \top$
                \State \Call{solveAbstract}{$\langle I, k \rangle, \emptyset$}
            \EndFor
        \EndFunction
        \algstore{u1}
    \end{algorithmic}

    \begin{algorithmic}
        \algrestore{u1}
        \Function{solveAbstract}{$\langle s, r \rangle , T$}
        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, T$} \Comment{Look for a candidate}
        \IIf{$r = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
        \State $T' \gets T$
        \Loop
            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, r \rangle, T, cand$} \Comment{Verify candidate}
            \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, T'$} \Comment{Solve refined AGT}
        \EndLoop
        \EndFunction
        \algstore{u2}
    \end{algorithmic}

    \begin{algorithmic}
        \algrestore{u2}
        \Function{findCandidate}{$\langle s, r \rangle, T$}
            \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
            \If{$sol = \texttt{unsat}$} 
                \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
                \State \Call{learn}{$\sigma, T$}
                \State \Return $\emptyset$ \Comment{No candidate exists}
            \Else
                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
            \EndIf
        \EndFunction
        \algstore{u3}
    \end{algorithmic}

    \begin{algorithmic}
        \algrestore{u3}
        \Function{verify}{$\langle s, r\rangle, gt, cand$}
            \For{$l \in leaves(gt)$}
            \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle s, r\rangle, gt, l$} \Comment{Get rank and state at leaf}
                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle r', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
                \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
            \EndFor
            \State \Return $\langle \True, \emptyset, \emptyset \rangle$
        \EndFunction
    \end{algorithmic}

    \caption{Unbounded Synthesis}
    \label{alg:unbounded}
\end{algorithm}

\subsection{Correctness}

The unbounded synthesis algorithm (Algorimthm \ref{alg:unbounded}) effectively
consists of two communicating solvers: the counterexample-guided bounded
reachability solver and the unbounded solver based on incremental induction.
Fortunately, correctness of both solvers can be established largely
independently, which simplifies the proof of correctness of the overall
algorithm.

We define two global invariants of the algorithm.  The \emph{may-invariant}
states that sets $B^m[i]$ grows monotonically with $i$ and that each $B^m[i+1]$
overapproximates the uncontrollable predecessor of $f^m[i]$: $$\forall
i<k.~B^m[i] \subseteq B^m[i+1], Upre(B^m[i]) \subseteq B^m[i+1].$$

The \emph{must-invariant} guarantees that the must-losing set $B^M$ is an
underapproximation of the actual losing set $B$: $$B^M \subseteq B$$.

The sets $B^m$ and $B^M$ are only modified by the inductive solver, implemented
by \textsc{learn} and \textoverline{\textsc{learn}} functions.  Below we prove that these
functions indeed maintain the invariants.

\subsection{Proof of \textsc{learn}}

We prove that postconditions of \textsc{learn} are satisfied assuming that its
preconditions hold.

Line~(4) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has
depth 1.  Consider formulas $f_1=\sigma(X_T) \land \textsc{treeFormula}(T_1)$
and $F_2 = \textsc{treeFormula}(T_2)$.  These formulas only share variables
$X_n$.  Their conjunction $F_1 \land F_2$ is unsatisfiable, as by construction
any solution of $F_1 \land F_2$ also satisfies $\sigma(X_T) \land
\textsc{treeFormula}(T)$, which is unsatisfiable (precondition (b)).  Hence the
interpolation operation is defined for $F_1$ and $F_2$.

Intuitively, the interpolant computed in line~(5) overapproximates the set of 
states reachable from $\sigma$ by following the tree from the root node to $n$,
and underapproximates the set of states from which the controller loses against 
tree $T_2$.

Formally, $\II$ has the property $\II \land F_2 \equiv \bot$.  Since $T_2$ is of depth
1, this means that the controller cannot force the
game into $W^m[|rank|(n-1)]$ when playing against counterexample moves in $T_2$.  Hence, 
$\II \cap Cpre(W^m[|rank|(n)]) = \emptyset$.  Furthermore, since the
may-invariant holds (assumption (c)), $\II \cap Cpre(i) = \emptyset$, for
all $i<|rank|(n)$.  Hence, removing $\II$ from all $W^m[i], i\leq |rank|(n)$
in line~(6) preserves the may-invariant, thus satisfying the first post-condition.

Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any assignment to 
$X_n$ that satisfies $\sigma(X_T) \land |treeFormula|(T_1)$ also satisfies $\II$.  Hence, removing $\II$ 
from $W_m[|rank|(n)]$ makes $\sigma(X_T) \land |treeFormula|(T_1)$ unsatisfiable, 
and hence all preconditions of the recursive invocation of \textsf{learn} in line~(7) 
are satisfied.

At the second las recursive call to \textsf{learn}, tree $T_1$ is empty, $n$ is the root node,
$|treeFormula|(T_1) \equiv W^m[|rank|(T)](X^T)$; hence 
$\sigma(X_T) \land |treeFormula|(T_1) \equiv \sigma(X_T) \land W^m[|rank|(T)](X^T) \equiv \bot$.
Thus the second postcondition of \textsf{learn} holds.

@@

We must prove that $|checkRank|$ terminates and that upon 
termination its postcondition holds, i.e., state $s$ is added to $W^M$ if 
its rank is no greater than $k$ or is removed from $W^m[k]$ otherwise.  
Termination follows from completeness of couterexample-guided search, which 
terminates after enumerating all possible opponent moves in the worst case.

Assume that $|rank|(s) > k$.  This means that at some point the algorithm
discovers a counterexample tree of rank $k$ that is losing for the controller 
from state $s$.  The algorithm then invokes the $|learn|$ method, which 
eliminates $s$ from $W^m[k]$.  Alternatively, if $|rank|(s) \leq k$, a 
counterexample tree losing for the environment will be discovered and 
subsequent call to $\overline{|learn|}$ adds $s$ to $W^M$.

\section{Optimisations}

\subsection{Generalising the initial state}

\subsection{Choosing sensible opponent moves}
